---
title: "Implementating and testing the HAR-QREG model i R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this R Markdown document we will implement the HAR-QREG model of [Haugom et al. (2016)](http://www.sciencedirect.com/science/article/pii/S1544612315001385) in R, and backtest the quantile/Value at Risk (VaR) predictions.
Predicting VaR is essentially predicting quantiles and hence quantile regression is a natural choice when modelling VaR. The HAR-QREG model is a quantile regression model using short term, medium term and long term historical volatility (square root of 1 day, 5 days and 20 days average squared returns) as explanatory variables. This is of course heavily inspired by the HAR-RV model of [Corsi (2009)](https://www.scopus.com/record/display.uri?eid=2-s2.0-62849101579&origin=inward&txGid=a69b94b190edf2833685a89bf157a243).

Bellow we will predict the 5 % quantile (VaR) on daily returns from the DAX index, using a rolling window of 1000 days.

## Code
For implementing and testing the model we only need to load/install two packages. The [GAS](https://www.rdocumentation.org/packages/GAS/versions/0.2.4) package has some nice financial data in addition to implementation of the [Conditional Coverage test](https://www.jstor.org/stable/2527341?seq=1#page_scan_tab_contents) and the [Dynamic Quantile test](http://www.simonemanganelli.org/Simone/Research_files/caviarPublished.pdf). And of course the [quantreg](https://cran.r-project.org/web/packages/quantreg/index.html) package for quantile regression.

```{r message = FALSE}
library(GAS)
library(quantreg)
```

Loading the data set *StockIndices*, from the GAS package, which includes almost 10 years of daily data for three European Stock indices.

```{r}
data("StockIndices")
head(StockIndices)
```

We will continue with the DAX index in this example and proceed with creating the three explanatory variables with use of the [filter function](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/filter.html). Notice that by removing the last observation and include an *NA* in the beginning of the vectors, we get lagged observations. This means that the volatility estimates on a given row, is not estimated using the actual return on that row.

```{r}
y <- StockIndices[, "DAX"]
n <- length(y)

# Make data frame with explanatory variabels
df1 <- data.frame(y = y,
                  sigma1 = c( NA, sqrt(y^2)[-n] ), # Square root of yesterdays squared return
                  sigma5 = c( NA, sqrt(filter(y^2, rep(1/5, 5), sides = 1))[-n] ), # Square root of past 5 days average squared return
                  sigma20 = c( NA, sqrt(filter(y^2, rep(1/20, 20), sides = 1))[-n] ) # Square root of past 20 days average squared return
)

head(df1, 25)
```

After removing the rows with missing observations we may estimate and predict 5% VaR with a rolling window of 1000 trading days within a for-loop. For comparison we also include the empirical 5% quantile of the 1000-day window (often referred to as historical simulation). 

```{r}
df1 <- na.omit(df1)
N <- nrow(df1)
# Rolling window one-day-ahead predictions with window of 1000 observations
win_size <- 1000 # Window size
pred0 <- numeric(N - win_size) # Historical quantiles
pred1 <- numeric(N - win_size) # HAR-QREG model
alpha <- 0.05 # Quantile to predict

for(r in win_size:(N -1)){
  rqfit <- rq(y ~ sigma1 + sigma5 + sigma20, data = df1[(r - win_size + 1):r, ], tau = alpha)
  
  pred0[r - win_size + 1] <- quantile(df1$y[(r - win_size + 1):r], probs = alpha)
  pred1[r - win_size + 1] <- predict(rqfit, df1[r + 1, ])
}

# A matrix with asset return and predicted quantiles
predmat <- cbind(Return = y[(win_size + 1):N], HS = pred0, HARQREG = pred1)
head(predmat)
```

Above is the first row of a matrix containing the asset return together with the 5% VaR predictions of the Historical Simulation model and the HAR-QREG model. The historical quantiles are, as expected, more static in nature than the predictions from the HAR-QREG model, which responds to the ever changing volatility. This might also be depicted in a plot.

```{r}
plot(as.Date(rownames(predmat)), predmat[, "Return"], type = "l", ylab = "Retun in %", xlab = "Date")
lines(as.Date(rownames(predmat)), predmat[, "HS"], col = 2)
lines(as.Date(rownames(predmat)), predmat[, "HARQREG"], col = 3)
```

The red line is the historical 5% quantile and the green line is the predicted 5% VaR from the HAR-QREG model.

Finally we can use the [BacktestVaR function](https://www.rdocumentation.org/packages/GAS/versions/0.1.9/topics/BacktestVaR) to evaluate the predictions. We extract the P-values from the Conditional Coverage test and the Dynamic Quantile test. In both this tests the null-hypothesis is that the predicted quantiles are correct and that there is no dependence in whether the asset returns is above/bellow the predicted quantile. Hence a larger P-value gives us more confidence in the model. 

Bellow the P-values are presented with a bar plot.
```{r}
btest0 <- BacktestVaR(data = predmat[, "Return"], VaR = predmat[, "HS"], alpha = alpha)
btest1 <- BacktestVaR(data = predmat[, "Return"], VaR = predmat[, "HARQREG"], alpha = alpha)

# A grafical presentation of P-values for Conditional Covareage test and Dynamic Quantile test
pval0 <- c("CC" = unname(btest0$LRcc[2]), "DQ" = btest0$DQ$pvalue)
pval1 <- c("CC" = unname(btest1$LRcc[2]), "DQ" = btest1$DQ$pvalue)

barplot(c(pval0, pval1), ylim = c(0, 1), col = c(2,2,3,3), main = "P-values")
abline(h = 0.05) # Illustrating a 5 % significance level
legend("topleft", legend = c("Historical Simulation", "HAR-QREG"), text.col = c(2,3),
       box.lty = 0)

```

